A quantitative framework for investigating the reliability of network construction

### 2018-03-19 manuscript/CI_interactions.tex
converted bib to Ecol Lett style

  Descriptions of ecological networks typically assume that the same interspecific interactions occur wherever and whenever a community is observed. This contrasts with the known stochasticity of ecological communities: community composition, species abundances, and link structure all vary in space and time. On top of this, finite sampling will generate variation in the set of interactions that are actually observed. Despite recognising these challenges, we lack the quantitative methodology to deal with the uncertainty generated by “true” spatiotemporal variation in ecological interactions or by incomplete sampling. Here we develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. To define the problem, we identify the different contributions to the uncertainty of an interaction and discuss the implications of this uncertainty in the estimation of network properties. We then develop a framework to explicitly quantify the uncertainty around each interaction. We illustrate this framework using the most extensively sampled network to date, demonstrating the universal uncertainty in network studies. Finally, we offer tangible recommendations for improving descriptions of ecological networks and for better sampling of rare interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

    In this study, we formalize the description of interactions between species as probabilities and develop analytical tools to capture the uncertainty in the estimation of these interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions [[double-check that we do this]]. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

      % %DG: the following subsection could be moved to the appendix if we need to cut down
      % Add to appendix.

      %   \subsection*{Computing the credible interval around a probability estimate}

      %       As well as obtaining posterior means and variances, it is possible to calculate a 95\% credible interval about the posterior mean. This is again straightforward to do in R (see the function ``credible\_interval" in \emph{Appendix S2}). Calculating the 95\% credible intervals for species pairs with 0 to 50 observed co-occurrences without any observed interaction (as in Fig.~\ref{Salix_pdfs}), we see that the interval narrows dramatically as the number of observations increases. In no case, however, does the confidence interval include 0. Even for a pair of species which appeared together in all 374 sites but were never observed interacting, the confidence interval (0.001, 0.018) comes close to, but does not include zero (Table~\ref{interval_table}). In other words, even with a very large data set, we will have problems in establishing that a given species pair can never interact.

      %       % # R code to make the table:
      %       % output=matrix(nrow=length(ns),ncol=5)
      %       % ns=c(0,1,2,5,10,15,20,25,50,100,150,200,374)
      %       % output[,1]=ns
      %       % for(r in 1:nrow(output)){
      %       %   interval1=credible_interval(calculate_parameters(sg_int_probs,ns[r],0),0.025,0.975)    
      %       %   interval2=credible_interval(calculate_parameters(gp_int_probs,ns[r],0),0.025,0.975)    
      %       %   output[r,2:3]=interval1
      %       %   output[r,4:5]=interval2
      %       % }

      %       \begin{table}[h]
      %         \caption{95\% credible intervals for the interaction probability of two species that co-occur in $N$ sites but are never observed to interact, obtained using prior information from the \emph{Salix}-galler network in~\citet{Barbour2016,Barbour2016Dryad}. [Tomas notes that this is conceptually very similar to Fig.~\ref{Salix_pdfs}]}
      %         \label{interval_table}
      %         \begin{tabular}{l | c c | c c }
      %         \multirow{2}{*}{N} & \multicolumn{2}{c}{\emph{Salix}-galler} & \multicolumn{2}{c}{galler-natural enemy}\\
      %          & Lower bound & Upper bound \\
      %         \hline
      %         0   & 0.152 & 0.931 & 0.008 & 0.364 \\
      %         1   & 0.110 & 0.843 & 0.007 & 0.337 \\
      %         2   & 0.087 & 0.760 & 0.006 & 0.313 \\
      %         5   & 0.053 & 0.573 & 0.005 & 0.258 \\
      %         10  & 0.033 & 0.400 & 0.004 & 0.200 \\
      %         15  & 0.023 & 0.306 & 0.003 & 0.163 \\
      %         20  & 0.018 & 0.247 & 0.003 & 0.137 \\
      %         25  & 0.015 & 0.208 & 0.002 & 0.119 \\
      %         50  & 0.008 & 0.115 & 0.001 & 0.071 \\
      %         100 & 0.004 & 0.061 & 0.001 & 0.039 \\
      %         150 & 0.003 & 0.041 & \textless0.001 & 0.027 \\
      %         200 & 0.002 & 0.031 & \textless0.001 & 0.021 \\
      %         \hline
      %         374 & 0.001 & 0.017 & \textless0.001 & 0.011 \\
      %         \hline
      %         \end{tabular}
      %         \end{table}

  \subsection*{How many samples are required to reach a minimal precision?}

@@

    \bibliographystyle{ecollett} 
    \bibliography{MyCollection} % Abbreviate journal titles.

>>>>

  Descriptions of ecological networks typically assume that the same interspecific interactions occur wherever and whenever a community is observed. This contrasts with the known stochasticity of ecological communities: community composition, species abundances, and link structure all vary in space and time. Moreover, finite sampling generates variation in the set of interactions actually observed. Here we develop the conceptual and analytical tools needed to capture uncertainty in the estimation of pairwise interactions. To define the problem, we identify the different contributions to the uncertainty of an interaction and its implications for the estimation of network properties. We then outline a framework to quantify the uncertainty around each interaction. We illustrate this framework using the most extensively sampled network to date. Finally, we offer both code and tangible recommendations for improving descriptions of ecological networks and sampling of rare interactions. Through these efforts, we demonstrate the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies. Most importantly, we stress that networks are best thought of as systems constructed from random variables, the stochastic nature of which must be acknowledged for an accurate representation. Doing so will fundamentally change networks analyses, but should lead to greater realism.

@@

    In this study, we formalize the description of interactions between species as probabilities and develop analytical tools to capture the uncertainty in the estimation of these interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first identify the different contributions to the uncertainty of an interaction and discuss the implications of each source of uncertainty for the properties of ecological networks. Next, we develop an analytical framework to quantify the uncertainty around interactions in an empirical web. We illustrate this framework using the most extensively sampled network to date (Box 1).  Finally, we offer tangible recommendations for improved descriptors of ecological interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

  \subsection*{How many samples are required to reach a minimal precision}

@@

    \bibliographystyle{ecol_let} 
    \bibliography{manual_abbrev} % Abbreviate journal titles.
